Community biomass and species richness temporal trends reveal contrasting

### 2022-02-22 paper/manuscript.Rmd
Update figures

@@

Current knowledge on the links among different facets of biodiversity mainly
comes from from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016 @danet_species_nodate]. space-for-time substitution began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time, i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the non-compliance of space-for-time substitution
[@adler_contrasting_2007; @blois_space_2013].

@@

The documentation of biodiversity temporal trends generally neglect species
interactions although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; @oconnor_general_2016]. In food-webs, spatial
comparisons in empirical and theoretical studies generally reported that trophic
level positively scales to community biomass [@wang_biodiversity_2018;
@danet_species_nodate]. Trophic level has also been shown to stabilise the
temporal variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012]. Our knowledge are even more scarce on the variability of
food-web structure in time and particularly how those temporal variations are linked to
the temporal variations of other aspects of biodiversity such as species richness
and community biomass. Spatially, connectance has been shown to be negatively
related to species richness in empirical food-webs [reviewed in
@dunne_network_2006]. Connectance has been shown to scale positively or
negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness temporal trends are associated to
temporal trends in food-web structure.

@@

@@

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure, evenness of biomass  and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)), $R_{cond}^2$ being the $R^2$conditional to the random effect. 
---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

@@

@@

@@

@@

## The temporal of community biomass and species richness relates on different assembly and disassembly rules

@@

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time in a given site
when community biomass increased through time. Reversely, connectance decreased
in food-webs when community biomass decreased through time. It suggests that
community biomass losses and gains are associated to losses and gains in large
bodied predators. This interpretation is in line with the results that community
biomass losses through time were associated to a temporal decrease of average
trophic level and that community biomass temporal gains were associated to
temporal increase of average trophic level. Another piece of evidence is that
evenness of biomass, measured by the Pielou index, decreased through time as
community biomass increased, meaning that biomass distribution across species
becomes less even when community biomass increased. Reversely, in food-webs that
displayed temporal decrease in community biomass, we found that evenness of
biomass increased, i.e. that biomass distribution across species becomes more
even.  Because predator-prey interactions relate in part on bodysize and thus
bodymass to a certain extent, we may expect that a given gain in predators of
higher trophic level leads to a decrease of biomass evenness and reversely. In
the same line, we found that the temporal trends of community biomass were positively
correlated to the temporal trends of the number and proportion of piscivorous
trophic species present in fish communities (Fig. SXX).

In contrast, we found that temporal trends of connectance and average trophic level
were not related to the temporal trends of species richness. However, we found
that temporal trends of the number of piscivorous species and piscivorous
trophic species were strongly positively linked to the temporal trends of
species richness (Fig. S3). It suggests that the temporal decrease of species
richness observed in our data are concomimant to losses of
piscivorous and non-piscivorous species, leading to the absence of temporal
trends in connectance and average trophic level and symmetrically for the
temporal increase of species richness. Furthemore, we found that the temporal
trends of biomass evenness had a positive relationship with the one of species
richness, in contraction of what we found for temporal trends of community biomass.
Together with the fact that species richness temporal trends had a positive
relationship with the temporal trends of community biomass, it suggests that
positive and negative  temporal trends of species richness are linked to gains and
losses of redundancy in food-webs, without overall effects on connectance
and average trophic level.

@@

In time and along spatial gradient, we found that connectance was positively
linked to community biomass. This result is also in accordance with theoretical
work showing that higher food-web connectance can result in higher community
biomass [@thebault_food-web_2003], but other display negative relationship
[@poisot_trophic_2013], highlighting that the relation between connectance and
community biomass may be dependant of assembly processes at work.]. In the same
line, we found that average trophic level had a positive relationship with
community biomass in space and time. It is in accordance with a theoretical
model showing that the presence of higher predators can lead to higher community
biomass [@wang_biodiversity_2018].  We also found in space and time that
evenness of biomass was negatively related to community biomass. This result is
in accordance with the observation of top-heavy shape of biomass pyramids in
food-webs, particularly in freshwater systems [@mccauley_prevalence_2018]. Many
mechanisms can be responsible have been invoked to explain this pattern but
omnivory, which is present in most of piscivorous species in our dataset, may
favor this shape by increasing food transfer efficiency toward higher trophic
level.

Surprisingly, we found that the temporal trends of connectance was not related
to the one of species richness as found spatially in most empirical food-web
[@dunne_network_2006] and also in the present study. We explained this pattern
by the fact that the temporal variation of species richness in our dataset seems
to be linked to variation in the level of redundancy in the role in the
food-webs rather than overall change of the vertical structure of food-webs. The
same rational can applied with average trophic level. *This discrependacy might
be due to the crucial assumptions that variations in space will correspond to
differences of assembly stages and then reflect the assembly/disassembly
trajectories in time of communities. Temporal variation of species richness
might relate on transitory dynamics while spatial gradient reflect aggregated
dynamics over longer time scale including a far number of processes including
high species turnover along spatial gradient.* This result is in line with other
studies which found mismatch between spatial and temporal patterns
[@adler_contrasting_2007; @johnson_testing_2008].  
However, we found agreement between space and time analysis for the positive
relationship between evenness of biomass and species richness. Another reason
might be that the spatial gradient displays a higher range of variation for
species richness than the temporal stations (Fig. S2). However, it might not
explain why we found strong relationships between the temporal trend of species
richness from one side and the temporal trends of community biomass, biomass
evenness, the number of piscivore species (Fig. S3). *No more idea, need
coffee*.

However, our result that food-web structure does not change a lot with temporal
variation of species richness is in line with the insurance hypothesis
[@yachi_biodiversity_1999]. In this view, the temporal loss and gain of species
richness represent respectively loss and gain of redundancy in the
food-web structure. Our results have profound implications for conservation. As
more and more studies reported community biomass and species richness trends
[@hallmann_more_2017; @ceballos_biological_2017], we documented that they might
not be equivalent in terms of consequences on the structure of the community. In
particular, the loss of the highest trophic level associated to the loss of
community biomass could have dramatic consequences on ecosystem functioning
as the highest predators are prone to stabilise communities
[@shanafelt_stability_2018; @danet_species_nodate]. In contrasts, the loss of
species richness might be linked to loss of redundancy in the food-web,
highlighting that those communities may become more at risk regarding future
perturbations [@yachi_biodiversity_1999]. Overall, it suggests that monitoring
several facets of biodiversity is essential to infer how ecological communities
are reshaped by loss and gain of biomass and species richness.

We further found that most temporal trends of food-webs were classified as
stable for species richness and community biomass. We found almost the same
number of stations that displayed temporal decrease or increase for species
richness, and twice more stations displayed decrease in community biomass than
increase. While, the spatial coverage of our data is good, it would be rather
risky to conclude about biodiversity trends in freshwater fish communities in
France. Mainly because we did not assess pressures that may affect
the temporal trends such as land-use changes and species invasions
[@comte_climate_2021; @su_human_2021].

@@

Our study suggests that richness and biomass temporal trends in natural fish
communities relate on different ecological processes, biomass temporal dynamics
being related to variations of at highest trophic levels while species richness
impacts redundancy of food-web structure. Our result also suggests that space
for time substitution may be relevant to predict the effects of community
biomass losses or gains on food-web structure. However, it might not the case
for species richness temporal variation. We look forward to next investigations
to see the generality of assembly rules in food-webs. There are crucial to
better understand the temporal dynamics of food-web structure and then its
impact on ecosystem functioning.

@@

# Figures {-}

```{r}
fig_mat_cap <- paste0(
  "(A) Distribution of the fishing sites. (B-F) Temporal trends of biomass,
  species richness, connectance, average trophic level and evenness of biomass
  for two stations. (À retravailler)"
)
```
```{r matmet, fig.cap=fig_mat_cap, out.width="100%"}
knitr::include_graphics(get_mypath("figures", "mat_met_inkscape.pdf"))
```

```{r}
fig_bm_cap <- paste0(
  "Relationship between community structure and community biomass. Left panel
  displays the relationships between the temporal trends of community structure
  and the one of community biomass. Lines correspond to the predictions of the
  linear models while the points are colored according the median community
  biomass of the three first years of the given temporal series. The lines
  surrounding each point represent the standard error associated to the temporal
  trends. Right panel displays the
  spatial relationship between median community structure over time and the
  median community biomass over time."

)

```{r}
fig_bef_cap <- paste0(
  "Relationship between the temporal trends of community structure
  and the one of species richness. Lines correspond to the predictions of the
  linear models. The lines surrounding each point represent the standard error
  associated to the temporal trends."
)
```
```{r bef, fig.cap=fig_bef_cap}
library(ggExtra)
ggMarginal(fig1_2$rich$log_bm_std, type = "histogram")
```

```{r}
fig_bm_rich_cap <- paste0(
  "Relationship among the temporal trends of community structure
  and the one of species richness and community biomass. The lines
  surrounding each point represent the standard error associated to the temporal
  trends."
)
```
```{r bmr, out.width = "100%", fig.cap=fig_bm_rich_cap}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~ .x +
    theme(axis.text = element_text(size = 6),
    axis.title = element_text(size = 8))
  )
fig2$bm <- map(fig2$bm,
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    )
)

# Add histograms
fig2$bm <- map(fig2$bm,~ggMarginal(.x, type = "histogram"))
fig2$rich <- map(fig2$rich,~ggMarginal(.x, type = "histogram"))

plot_grid(
  plot_grid(plotlist = fig2$bm, nrow = 1, labels = c("A", "B", "C")),
  plot_grid(plotlist = fig2$rich, nrow = 1, labels = c("D", "E", "F")),
  nrow = 2
)
```

```{r bm, eval=FALSE, fig.height = 12, fig.cap=fig_bm_cap}
ti <- list()
for (
  i in c("log_rich_std", "w_trph_lvl_avg", "piel_bm", "piel_nind", "ct_ff")
  ) {
  ti[[i]] <- list(p_bm[[i]], p_sp_bm[[i]])
}

# Make the label on two lines
fig1 <- plot_grid(plotlist = c(ti[[1]], ti[[2]], ti[[3]], ti[[4]], ti[[5]]), ncol = 2)
fig1
```

```{r}
p_rich <- fig1_2$rich
p_sp_rich <- sp_fig1_2_bm_rich_mono_stable_trends$rich
```

```{r rich, eval=FALSE,fig.height = 12, fig.cap=fig_rich_cap}
trich <- list()
for (i in c("log_bm_std", "w_trph_lvl_avg",  "piel_bm", "piel_nind", "ct_ff")) {
  trich[[i]] <- list(p_rich[[i]], p_sp_rich[[i]])
}

# Make the label on two lines
fig2 <- plot_grid(
  plotlist = c(trich[[1]], trich[[2]], trich[[3]], trich[[4]], trich[[5]]),
  ncol = 2)
fig2
```

```{r}
std_cap <- paste0(
  "Standardized coefficients of the effects of community biomass and species
  richness on food-web structure for the temporal and spatial models. The lines
  display the standard errors."
)
```

```{r std, fig.cap=std_cap}
loadd(fig_std_coef)
fig_std_coef
```

@@

```{r}
loadd(reg_table_bm_rich_mono_stable_trends)
reg_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
    ) %>%
kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression coefficients for the linear models explaining the links
      between the temporal trends of community biomass and the temporal trends
      of the network structure."),
      label = "reg"
      ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
rsq_table_bm_rich_mono_stable_trends %>%
  mutate(
    Response = str_replace_all(Response, var_replacement())
  ) %>%
  kable(., booktabs = TRUE,
    caption =
      paste0(
      "Marginal and conditional squared R for temporal models."),
  label = "rsq"
      ) %>% #
  kable_styling(latex_options = c("striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

>>>>

```{r}
loadd(rsq_table_bm_rich_mono_stable_trends)
loadd(reg_table_bm_rich_mono_stable_trends)
```

@@

Current knowledge on the links among different facets of biodiversity mainly
comes from from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016; @danet_species_nodate]. space-for-time substitution began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time, i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the non-compliance of space-for-time substitution
[@adler_contrasting_2007; @blois_space_2013].

@@

The documentation of biodiversity temporal trends generally neglect species
interactions although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; @oconnor_general_2017]. In food-webs, spatial
comparisons in empirical and theoretical studies generally reported that trophic
level positively scales to community biomass [@wang_biodiversity_2018;
@danet_species_nodate]. Trophic level has also been shown to stabilise the
temporal variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012]. Our knowledge are even more scarce on the variability of
food-web structure in time and particularly how those temporal variations are linked to
the temporal variations of other aspects of biodiversity such as species richness
and community biomass. Spatially, connectance has been shown to be negatively
related to species richness in empirical food-webs [reviewed in
@dunne_network_2006]. Connectance has been shown to scale positively or
negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness temporal trends are associated to
temporal trends in food-web structure.

@@

```{r}

fig_mat_cap <- paste0( "Using  sites of stream fishes in French streams that
  have been monitored at least 10 years, 1) we infered the food-web at each site
and each sampling event based on ontogenic food diet information for each
species and predator-prey body size ratio. 2) We computed the temporal trends
and the median of community biomass, species richness, connectance, average
trophic level and evenness of biomass at each site. 3) We then assessed the
relationships among a) temporal trends of food-web structure. We also assessed
the spatial relationships among median food-web structure while controlling for
the environmental gradient with two axis of a rotated PCA. The first axis of the
PCA is related to the stream gradient from upstream to downstream while the
second is related to enrichment (BOD) and site altitude (error on the PCA,
  station is present!!!).")

```

```{r matmet, fig.cap=fig_mat_cap, out.width="80%"}
knitr::include_graphics(get_mypath("figures", "mat_met_inkscape.pdf"))
```

@@

@@

Overall, we found high heterogeneity in the explicative power of statistical
models linking temporal dynamics of food-web structure, evenness of biomass  and the one of community
biomass and species richness spanning from $R_{adj}^2= `r rsq_piel_nind`$ for
the evenness of abundances to $R_{adj}^2= `r rsq_tlvl`$ for average trophic
level (Table \@ref(tab:rsq)). The models over the spatial gradient were more performant, with
$R_{cond}^2$ varing from $`r rsq_tlvl_sp`$ to $`r rsq_piel_bm_sp`$ (Table
\@ref(tab:rsq-sp)), $R_{cond}^2$ being the $R^2$conditional to the random effect.

---
#Overall, we found no significant effect of the temporal gain or losses of
#communinity biomass biomass and species richness on the relationship between the
#temporal trends of community structure and that of biomass and species richness
#(Table \@ref(tab:aov)).
---

```{r}
fig_bef_cap <- paste0(
  "Relationship between the temporal trends of community biomass 
  and the one of species richness. Red line correspond to the predictions of the
  linear models. The lines surrounding points display the standard error
  associated to the temporal trends. The histograms display the distribution of
  response and predictive variables."
)
```

```{r}
library(ggExtra)
coeff_bef <- c(
  rsq = rsq_table_bm_rich_mono_stable_trends %>%
    filter(Response == "log_bm_std") %$%
    unlist(.[, c("R.squared")]),
  reg_table_bm_rich_mono_stable_trends %>%
    filter(response == "log_bm_std", term == "log_rich_std") %$%
    unlist(.[, c("estimate", "std.error", "std_estimate", "stdse")])
  ) %>%
round(., 2)

beta_rsq <- function(
  rsq = NULL,
  beta = NULL,
  se = NULL) {

  tmp <- "paste(
  beta, \" (s.e) \", \" = \",
  mybeta, \" (\", myse, \"), \",

  italic(R)^2, \" = \", myrsq
  )"

str_replace_all(tmp,
  c(
    "myrsq" =  as.character(rsq),
    "mybeta" = as.character(beta),
    "myse" = as.character(se)
  )
)

}
```

```{r bef, fig.cap=fig_bef_cap, out.width="50%"}
p_bef <- fig1_2$rich$log_bm_std +
  annotate("text", x = -.05, y = .3, parse = T,
    label = beta_rsq(rsq = coeff_bef["rsq"],
      beta = coeff_bef["estimate"], se =
        coeff_bef["std.error"])
   ) +
  labs(
    y = "Log community biomass (g per year)",
    x = "Log species richness (per year)"
  )

ggMarginal(p_bef, type = "histogram")
```

@@

```{r}
fig_bm_cap <- paste0(
  "Relationship between community structure and community biomass. Left panel
  displays the relationships between the temporal trends of community structure
  and the one of community biomass. Lines correspond to the predictions of the
  linear models while the points are colored according the median community
  biomass of the three first years of the given temporal series. The lines
  surrounding each point represent the standard error associated to the temporal
  trends. Right panel displays the
  spatial relationship between median community structure over time and the
  median community biomass over time."
)

```{r}
fig_bm_rich_cap <- paste0(
  "Relationships between the temporal trends of food-web structure
  and the ones of community biomass (left panel) species richness (right panel).
  The lines surrounding each point display the standard errors associated to
  the estimated temporal trends. Red and black lines represent respectively
  significant relationships. The histograms display the distribution of
  response and predictive variables.")
```

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 6),
      axis.title = element_text(size = 8)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = paste0("Log biomass",
      "\ntrends (per year)"),
  )
)
# Delete redondant legend
fig2$bm[1:2] <- map(fig2$bm[1:2],
  ~ .x + theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )
  )

fig2$rich <- map(fig2$rich,
  ~ .x + theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  labs(x = paste0("Log species richness",
      "\ntrends (per year)"))
  )
fig2$rich[1:2] <- map(fig2$rich[1:2],
  ~ .x + theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )
  )

# Add histograms
fig2$bm[[1]] <- ggMarginal(fig2$bm[[1]], type = "histogram", margins = "x")
fig2$rich[[1]] <- ggMarginal(fig2$rich[[1]], type = "histogram", margins = "both")
fig2$rich[2:3] <- map(fig2$rich[2:3], ~ggMarginal(.x, type = "histogram", margins = "y"))

```

```{r bmr, out.width="70%", fig.height=7, fig.cap=fig_bm_rich_cap}
plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("A", "B", "C"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("D", "E", "F"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(1, 1)
)
```

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Effect of community biomass (top
  panel) and species richness (bottom panel) on food-web structure for the
temporal and spatial models. Top of
  bars display the standardized coefficients from the linear models, the lines
  around display the standard errors.")
```

```{r std, fig.cap=std_cap, fig.width=6}
#loadd(fig_std_coef)
std_table_bm_rich_mono_stable_trends %>%
  filter(
    term %in% c("log_rich_std", "log_bm_std"),
    !response %in% "piel_nind") %>%
  mutate_at(vars(term, response), ~str_replace_all(., var_replacement(
        several_lines = TRUE))) %>%
  mutate(
    type = str_to_sentence(type),
    type = factor(type, levels = c("Temporal", "Spatial"))) %>%
  ggplot(aes(y = std_estimate, x = response, fill = type)) +
  viridis::scale_fill_viridis(begin = 0.5, end = 1, discrete = TRUE) +
  facet_grid(rows = vars(term)) +
  geom_col(position = position_dodge(width = .9)) +
  geom_linerange(aes(ymin = std_estimate - stdse, ymax = std_estimate + stdse),
    position = position_dodge(width = .9)) +
  labs(x = "Response variables", y = "Standardized slopes", fill = "Model") +
  theme(legend.position = c(0.7, .1),
    legend.direction = "horizontal",
    text = element_text(size = 8), axis.text = element_text(size = 8)) +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = "Predictive variable",
      breaks = NULL, labels = NULL))
```

@@

We further found that most temporal trends of food-webs were classified as
stable for species richness and community biomass. We found almost the same
number of stations that displayed temporal decrease or increase for species
richness, and twice more stations displayed decrease in community biomass than
increase. This result is in accordance with previous meta analysis on species
richness trends of vertebrate in marine, freshwater and terrestrial systems
[@dornelas_assemblage_2014; @blowes_geography_2019; @klink_meta-analysis_2020].
While the spatial coverage of our data is good, it would be rather risky to
conclude about biodiversity trends in freshwater fish communities because we did
not assess pressures that may affect the temporal trends such as land-use
changes and species invasions [@comte_climate_2021; @su_human_2021;
@cardinale_is_2018].

@@

## Food-web structure changes according to the temporal trends of community biomass and species richness

@@

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time. It suggests that
community biomass losses and gains are associated to losses and gains in large
bodied piscivores. This interpretation is in
line with the results that community biomass losses through time were associated
to a temporal decrease of average trophic level and that community biomass
temporal gains were associated to temporal increase of average trophic level.
Another piece of evidence is that evenness of biomass, measured by the Pielou
index, decreased through time as community biomass increased, meaning that
biomass distribution across species becomes less even when community biomass
increased. Reversely, in food-webs that displayed temporal decrease in community
biomass, we found that evenness of biomass increased, i.e. that biomass
distribution across species becomes more even. Because predator-prey
interactions relate in part on bodysize and thus bodymass to a certain extent,
we may expect that a given gain in predators of higher trophic level leads to a
decrease of biomass evenness and reversely. In the same line, we found that the
temporal trends of community biomass were positively accociated to the temporal
trends of the number and proportion of piscivorous trophic species present in
fish communities (Fig. SXX).

In contrast, we found that temporal trends of connectance and average trophic
level were not related to the temporal trends of species richness. However, we
found that temporal trends of the number of piscivorous species and piscivorous
trophic species were positively associated to the temporal trends of species
richness (Fig. S3). It suggests that the temporal decrease of species richness
observed in our data are concomimant to losses of piscivorous and
non-piscivorous species, leading to the absence of temporal trends in
connectance and average trophic level and symmetrically for the temporal
increase of species richness. Furthemore, we found that the temporal trends of
biomass evenness had a positive relationship with the one of species richness,
in contradiction of what we found for temporal trends of community biomass.
Together with the fact that species richness temporal trends had a positive
relationship with the temporal trends of community biomass, it suggests that
positive and negative  temporal trends of species richness are linked to gains
and losses of redundancy in food-webs, without overall effects on connectance
and average trophic level. This is in line with the insurance hypothesis
[@yachi_biodiversity_1999] which predicts the temporal loss and gain of species
richness represent respectively loss and gain of redundancy in the community,
here in the food-web structure.

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilise communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_nodate]. In contrasts, the loss of species richness might be
linked to loss of redundancy in the food-web, highlighting that those
communities may become more at risk regarding future perturbations
[@yachi_biodiversity_1999]. Overall, it suggests that monitoring several facets
of community structure is essential to infer how ecological communities are
reshaped by loss and gain of biomass and species richness.

@@

In time and space, we found that connectance was positively linked to community
biomass. This result is also in accordance with theoretical work showing that
higher food-web connectance can result in higher community biomass
[@thebault_food-web_2003], or lower [@poisot_trophic_2013], highlighting that
the relation between connectance and community biomass may be dependant on how
food-web structure change with community biomass. Likewise, we found that
average trophic level had a positive relationship with community biomass in
space and time, in accordance with a theoretical study predicting an increase of
community biomass subsequent to the presence of higher predators
[@wang_biodiversity_2018]. We also found that biomass evenness was negatively
related to community biomass in space and time. This result is in accordance
with the reported top-heavy shape of biomass pyramids in food-webs,
particularly in freshwater systems [@mccauley_prevalence_2018]. Many mechanisms
have been invoked to explain this pattern but omnivory, which
is present in most of piscivorous species in our dataset, may favor this shape
by increasing food transfer efficiency toward the higher trophic levels.

Surprisingly, we found that the temporal trends of connectance were not related
to the one of species richness as found spatially in most empirical food-webs
[@dunne_network_2006] and also here. We explained this pattern by the fact that
the temporal variation of species richness in our dataset seems to be linked to
temporal variation of redundancy level in the structure of food-webs rather than
overall change of their vertical structure. The same rational can applied with
average trophic level. This result is in line with other studies which found
mismatch between spatial and temporal patterns [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepency in space and time might be due to the
crucial assumption that variations in space reflect differences in assembly
stages of ecological communities and then mirrors their temporal
assembly/disassembly trajectories. Temporal variation of species richness might
relate on transitory dynamics while spatial gradient reflect aggregated dynamics
over longer time scale including a far number of processes including high
species turnover along spatial gradient. However, we fount that the positive
relationship between evenness of biomass and species richness were consistent in
space and time. Another reason might be that the spatial gradient displays a
higher range of variation for species richness than the temporal stations (Fig.
S2). However, we found strong relationships between the temporal trend of
species richness from one side and the temporal trends of community biomass,
biomass evenness, and the number of piscivore species from the other side (Fig. S3).

@@

Our study suggests that richness and biomass temporal trends in natural fish
communities relate on different ecological processes, biomass temporal dynamics
being related to variations of food-web structure at the higher trophic levels
while species richness impacts redundancy of species in food-web structure. Our
result also suggests that space for time substitution may be relevant to predict
the effects of community biomass losses or gains on food-web structure. However,
our findings suggest that space-for-time substitution might not work for species
richness. Our study calls for further investigations to support the generality of
assembly rules in food-webs. They are crucial to better understand the
assembly and disassembly of food-web and then predict its impact on ecosystem
functioning.

@@

@@

```{r}
reg_table_bm_rich_mono_stable_trends %>%
  mutate(
    response = str_replace_all(response, var_replacement()),
    term = str_replace_all(term, var_replacement())
    ) %>%
kable(., booktabs = TRUE,
  row.names = FALSE,
  caption =
    paste0(
      "Regression coefficients for the linear models explaining the links
      between the temporal trends of community biomass and the temporal trends
      of the network structure."),
      label = "reg"
      ) %>%
  kable_styling(latex_options = c("scale_down", "striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```

```{r}
table_rsq_tps <- rsq_table_bm_rich_mono_stable_trends %>%
  mutate(
    Response = str_replace_all(Response, var_replacement())
  )
table_rsq_tps %>%
  kable(., booktabs = TRUE,
    caption =
      paste0(
      "Marginal and conditional squared R for temporal models."),
  label = "rsq"
      ) %>% #
  kable_styling(latex_options = c("striped")) %>%
  collapse_rows(columns = 1, valign = "top")
```
